---
title: "Analysis"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r, echo=FALSE, include = FALSE}
library(MASS)
library(tidyverse)
library(mosaic)
library(readr)
library(ipumsr)
library(labelled)
library(dplyr)
library(ggplot2)
library(generalhoslem)


highered_00006 <- read_csv("/home/pelayo1/Stat_316_Project/highered_00006.csv")
usa_ddi <- read_ipums_ddi("/home/pelayo1/Stat_316_Project/highered_00006.xml")
ed_data <- read_ipums_micro(usa_ddi, verbose = FALSE)

ed_data <- ed_data %>% 
  filter(LFSTAT==1, JOBSATIS != 98) %>%
  mutate(GENDER = droplevels(as_factor(GENDER)), 
         RACETH= droplevels(as_factor(RACETH)), 
         DGRDG= droplevels(as_factor(DGRDG)), 
         LFSTAT= droplevels(as_factor(LFSTAT)), 
         HRSWKGR = droplevels(as_factor(HRSWKGR)), 
         JOBINS = droplevels(as_factor(JOBINS)),
         JOBPENS=droplevels.factor(JOBPENS),
         JOBPROFT=droplevels(as_factor(JOBPROFT),
        JOBVAC = droplevels(as_factor(JOBVAC)),
        OCEDRLP = droplevels(as_factor(OCEDRLP)), 
        NOCPRMG = droplevels(as_factor(NOCPRMG)),
        WAPRSM= droplevels(as_factor(WAPRSM)),
        SALARY = droplevels(as_factor(SALARY)), 
        JOBSATIS = droplevels(as_factor(JOBSATIS))))
ed_data <- ed_data %>% mutate(OCEDRLP = ifelse(OCEDRLP == 1,"Closely Related",
                                    ifelse(OCEDRLP == 2, "Somewhat Related",
                                    ifelse(OCEDRLP == 3, "Not Related",
                                    ifelse(OCEDRLP == 98, "Logical Skip", NA)))))
ed_data <- ed_data %>% mutate(NOCPRMG = ifelse(NOCPRMG == 1,"Computer and mathematical scientists ",
                                    ifelse(NOCPRMG == 2, "Biological, agricultural and other life scientists",
                                    ifelse(NOCPRMG == 3, "Physical and related scientists",
                                    ifelse(NOCPRMG == 4, "Social and related scientists",
                                    ifelse(NOCPRMG == 5, "Engineers",
                                    ifelse(NOCPRMG == 6, "Science and engineering related occupations ",
                                    ifelse(NOCPRMG == 7, "Non-science and engineering occupations ",
                                    ifelse(NOCPRMG == 98, "Logical Skip", NA)))))))))

ed_data <- ed_data %>% mutate(WAPRSM = ifelse(WAPRSM ==1, "Research and Development",
                                    ifelse(WAPRSM == 2, "Teaching",
                                    ifelse(WAPRSM == 3,"Management and Administration", 
                                    ifelse(NOCPRMG ==4, "Computer Applications", "Other")))))
ed_data <- ed_data %>%
  mutate(AGE23 = AGE - 23)

ed_data$Age10 <- ed_data$AGE23/10

```

#Looking at work related factors
Intercept: closely related, Biological, agricultural and other life scientists, management and administration
All of these variables have significant p-values. 
-> we include all work-related factors explanatory variables of interest

Model 1
```{r}
wr1<- polr(as.factor(JOBSATIS) ~ OCEDRLP + NOCPRMG + WAPRSM, data = ed_data, Hess=TRUE)
(wr1_ctable <- coef(summary(wr1)))
wr1_p <- pnorm(abs(wr1_ctable[, "t value"]), lower.tail = FALSE) * 2
(wr1_ctable <- cbind(wr1_ctable, "p value" = wr1_p))

```

#Looking at the Available Benefits
SALARY gives the model error so we have to scale it

jb1 includes all job benefits + salary in the model -> we see that all variables significant

Now we look at interactions: potential interactions from EDA include insurance: pension plan and insurance:paid vacation/sick leave. 
jb2 includes all job benefits in model + 2 interactions mentioned above -> both interactions are not significant

Is jb2 (with interaction insurance:vacation) is significantly better than jb1? No, based on LRT test

Model 2a
```{r}
jb1 <- polr(as.factor(JOBSATIS) ~ scale(SALARY) +JOBVAC + JOBINS + JOBPENS + JOBPROFT, data = ed_data, Hess=TRUE)
(ctable1 <- coef(summary(jb1)))
p1 <- pnorm(abs(ctable1[, "t value"]), lower.tail = FALSE) * 2
(ctable1 <- cbind(ctable1, "p value" = p1))
```

Model 2b
```{r}
jb2 <- polr(as.factor(JOBSATIS) ~ scale(SALARY) + JOBVAC+ JOBINS + JOBPENS + JOBPROFT + JOBINS:JOBPENS +JOBINS:JOBVAC, data = ed_data, Hess=TRUE)
(ctable2 <- coef(summary(jb2)))
p2 <- pnorm(abs(ctable2[, "t value"]), lower.tail = FALSE) * 2
(ctable2 <- cbind(ctable2, "p value" = p2))

anova(jb1, jb2)
#####
```

Model 2c
```{r}
jb3 <- polr(as.factor(JOBSATIS) ~ scale(SALARY)+JOBVAC+ JOBINS + JOBPENS + JOBPROFT + scale(SALARY):JOBVAC +scale(SALARY):JOBINS, data = ed_data, Hess=TRUE)
summary(jb3)
anova(jb1, jb3)

```

#Looking at the Demographic factors
dmf1 includes all demographic factors -> all significant

Model 3
```{r}
dmf1 <- polr(as.factor(JOBSATIS) ~ Age10 + GENDER +RACETH, data = ed_data, Hess=TRUE)
exp(-0.8476)/(1 +exp(-0.8476)) #Probability of very satisfied for white

(ctable.dmf1 <- coef(summary(jb1)))
p.dmf1 <- pnorm(abs(ctable.dmf1[, "t value"]), lower.tail = FALSE) * 2
(ctable.dmf1 <- cbind(ctable.dmf1, "p value" = p.dmf1))
```

#Combine 3 groups of factors
final1 combine the 3 small models above (basically all variables we looked at) -> all significant. One problem is that the model is huge

One idea for reducing the model: we can look at effect size. Although signifcant, effect size of pension plan, gender and age is small. 

final2 excludes pension plan, gender and age. However, both AIC and LRT favors final1

Model 4a
```{r}
final1 <- polr(as.factor(JOBSATIS) ~ OCEDRLP + NOCPRMG + WAPRSM + scale(SALARY) +JOBVAC + JOBINS + JOBPENS + JOBPROFT + Age10 + GENDER +RACETH, data = ed_data, Hess=TRUE)
exp(-coef(final1))
AIC(final1)
BIC(final1)
(ctable.final1 <- coef(summary(final1)))
p.final1 <- pnorm(abs(ctable.final1[, "t value"]), lower.tail = FALSE) * 2
(ctable.final1 <- cbind(ctable.final1, "p value" = p.final1))

#confint(final1)

```

Model 4b
```{r}
final2 <- polr(as.factor(JOBSATIS) ~ OCEDRLP + NOCPRMG + WAPRSM + scale(SALARY) +JOBVAC + JOBINS + JOBPROFT +RACETH, data = ed_data, Hess=TRUE)
AIC(final2)
BIC(final2)
(ctable.final2 <- coef(summary(final2)))
p.final2 <- pnorm(abs(ctable.final2[, "t value"]), lower.tail = FALSE) * 2
(ctable.final2 <- cbind(ctable.final1, "p value" = p.final2))

anova(final1, final2)

```

Model 4c
```{r}
final3 <- polr(as.factor(JOBSATIS) ~ OCEDRLP + NOCPRMG + WAPRSM + scale(SALARY) +JOBVAC + JOBINS + JOBPENS + JOBPROFT + Age10 + GENDER +RACETH + GENDER:scale(SALARY), data = ed_data, Hess=TRUE)
summary(final3)
```


Model 4d
```{r}
final4 <- polr(as.factor(JOBSATIS) ~ OCEDRLP + NOCPRMG + WAPRSM + scale(SALARY) +JOBVAC + JOBINS + JOBPENS + JOBPROFT + Age10 + GENDER +RACETH + GENDER:scale(SALARY), data = ed_data, Hess=TRUE)
summary(final4)
```

Model 4e
```{r}
final5 <- polr(as.factor(JOBSATIS) ~ OCEDRLP + NOCPRMG + WAPRSM + scale(SALARY) +JOBVAC + JOBINS + JOBPENS + JOBPROFT + Age10 + GENDER +RACETH, data = ed_data, Hess=TRUE)
summary(final5)
```

Model 4f
```{r}
ed_data_salary <- ed_data[ed_data$SALARY!=0,]
final6 <- polr(as.factor(JOBSATIS) ~ OCEDRLP + NOCPRMG + WAPRSM + scale(SALARY) +JOBVAC + JOBINS + JOBPENS + JOBPROFT + Age10 + GENDER +RACETH, data = ed_data_salary, Hess=TRUE)
summary(final6)
```