---
title: "EDA"
output:
  word_document: default
  pdf_document: default
---

```{r, echo=FALSE, include = FALSE}
library(tidyverse)
library(mosaic)
library(readr)
library(ipumsr)
library(labelled)
library(dplyr)
library(ggplot2)

highered_00006 <- read_csv("/home/pelayo1/Stat_316_Project/highered_00006.csv")
#View(highered_00006)
```

```{r, data cleaning, echo = FALSE}
usa_ddi <- read_ipums_ddi("/home/pelayo1/Stat_316_Project/highered_00006.xml")
ed_data <- read_ipums_micro(usa_ddi, verbose = FALSE)

ed_data <- ed_data %>% 
  filter(LFSTAT==1, JOBSATIS != 98) %>%
  mutate(GENDER = droplevels(as_factor(GENDER)), 
         RACETH= droplevels(as_factor(RACETH)), 
         DGRDG= droplevels(as_factor(DGRDG)), 
         LFSTAT= droplevels(as_factor(LFSTAT)), 
         HRSWKGR = droplevels(as_factor(HRSWKGR)), 
         JOBINS = droplevels(as_factor(JOBINS)),
         JOBPENS=droplevels.factor(JOBPENS),
         JOBPROFT=droplevels(as_factor(JOBPROFT),
        JOBVAC = droplevels(as_factor(JOBVAC)),
        OCEDRLP = droplevels(as_factor(OCEDRLP)), 
        NOCPRMG = droplevels(as_factor(NOCPRMG)),
        WAPRSM= droplevels(as_factor(WAPRSM)),
        SALARY = droplevels(as_factor(SALARY)), 
        JOBSATIS = droplevels(as_factor(JOBSATIS))))
ed_data <- ed_data %>% mutate(OCEDRLP = ifelse(OCEDRLP == 1,"Closely Related",
                                    ifelse(OCEDRLP == 2, "Somewhat Related",
                                    ifelse(OCEDRLP == 3, "Not Related",
                                    ifelse(OCEDRLP == 98, "Logical Skip", NA)))))
ed_data <- ed_data %>% mutate(NOCPRMG = ifelse(NOCPRMG == 1,"Computer and mathematical scientists ",
                                    ifelse(NOCPRMG == 2, "Biological, agricultural and other life scientists",
                                    ifelse(NOCPRMG == 3, "Physical and related scientists",
                                    ifelse(NOCPRMG == 4, "Social and related scientists",
                                    ifelse(NOCPRMG == 5, "Engineers",
                                    ifelse(NOCPRMG == 6, "Science and engineering related occupations ",
                                    ifelse(NOCPRMG == 7, "Non-science and engineering occupations ",
                                    ifelse(NOCPRMG == 98, "Logical Skip", NA)))))))))

ed_data <- ed_data %>% mutate(NOCPRMG_non = ifelse(NOCPRMG =="Computer and mathematical scientists ", "MSCS",
                                    ifelse(NOCPRMG == "Biological, agricultural and other life scientists", "Bio",
                                    ifelse(NOCPRMG == "Physical and related scientists", "Physics",
                                    ifelse(NOCPRMG =="Social and related scientists", "Social",
                                    ifelse(NOCPRMG =="Engineers", "Engineers",
                                    ifelse(NOCPRMG =="Science and engineering related occupations ", "S-Engineer-related", "NS-Engineer-related")))))))


```

Job satisfaction
```{r, EDA from AGE}
ed_data %>%
  count(JOBSATIS)
ggplot(data = ed_data) + 
  geom_bar(mapping = aes(x = JOBSATIS))
```

#####
Demographics
#####


Age 
```{r}
favstats(~AGE, data=ed_data)
ggplot(data=ed_data) + 
  geom_histogram(mapping = aes(x=AGE))
favstats(AGE~JOBSATIS, data=ed_data)
ggplot(data = ed_data) + 
  geom_boxplot(mapping = aes(x = as.factor(JOBSATIS), y=AGE))
```

Gender
```{r}
#Gender
prop.table(table(ed_data$GENDER))
prop.table(table(ed_data$GENDER, ed_data$JOBSATIS),1)
prop.table(table(ed_data$GENDER, ed_data$JOBSATIS),2)

gender.prop <- prop.table(table(ed_data$JOBSATIS, ed_data$GENDER))
mosaicplot(gender.prop, color=TRUE,main="Job satisfaction by gender")
```

Race
```{r}
prop.table(table(ed_data$RACETH))
prop.table(table(ed_data$RACETH, ed_data$JOBSATIS),1)
prop.table(table(ed_data$RACETH, ed_data$JOBSATIS),2)
race.prop <- prop.table(table(ed_data$JOBSATIS, ed_data$RACETH))
mosaicplot(race.prop, color=TRUE,main="Job satisfaction by race")
```


#####
Education
#####


Highest degree
```{r}
prop.table(table(ed_data$DGRDG))
prop.table(table(ed_data$DGRDG, ed_data$JOBSATIS),1)
prop.table(table(ed_data$DGRDG, ed_data$JOBSATIS),2)
degree.prop <- prop.table(table(ed_data$JOBSATIS, ed_data$DGRDG))
mosaicplot(degree.prop, color=TRUE,main="Job satisfaction by highest degree")
```


Satisfaction verses Primary Work Activity
```{r}
ed_data <- ed_data %>% mutate(WAPRSM_non = ifelse(WAPRSM ==1, "Research",
                                    ifelse(WAPRSM == 2, "Teaching",
                                    ifelse(WAPRSM == 3,"Management", 
                                    ifelse(WAPRSM ==4, "Computers", "Other")))))


waprsm <- prop.table(table(ed_data$JOBSATIS, ed_data$WAPRSM_non),1)
waprsm
mosaicplot(waprsm, color=TRUE,main="Satisfaction vs Primary Work Activity")
```


Job Related To Major 
```{r, EDA from Major Related}
g <- ggplot(ed_data, aes(JOBSATIS, fill= OCEDRLP)) + geom_bar()
g
plyr::count(ed_data, "OCEDRLP")
table(ed_data$JOBSATIS, ed_data$OCEDRLP)

```

Primary Job Work
```{r, Analysis NOCPRNG}
table(ed_data$NOCPRMG_non, ed_data$JOBSATIS)

prop.table(table(ed_data$NOCPRMG, ed_data$JOBSATIS), 2) #Divides by Column
prop.table(table(ed_data$NOCPRMG, ed_data$JOBSATIS),1) #Divides by Row
prop.table(table(ed_data$NOCPRMG, ed_data$JOBSATIS)) #Overall


prop.table(table(ed_data$NOCPRMG, ed_data$JOBSATIS)) #Divides by Column
mosaicplot(prop.table(table(ed_data$NOCPRMG, ed_data$JOBSATIS)), color=TRUE,main="Overall NOCPRMG vs Satisfaction Proportions")


```

Interaction Between Relationship to Major and Primary Job Work
```{r}
ed_data <- ed_data %>% mutate(OCEDRLP_non = ifelse(OCEDRLP =="Closely Related", "Close",
                                    ifelse(OCEDRLP == "Somewhat Related", "Somewhat", "Not")))

prop.table(table(ed_data$NOCPRMG_non, ed_data$OCEDRLP_non),2) #Divides by Column
prop.table(table(ed_data$NOCPRMG_non, ed_data$OCEDRLP_non),1)

mosaicplot(prop.table(table(ed_data$NOCPRMG_non, ed_data$OCEDRLP_non)), color=TRUE,main="Overall NOCPRMG vs OCEDRLP Proportions")

```


####
Work Related
####


Hours/week worked
```{r}
prop.table(table(ed_data$HRSWKGR))
prop.table(table(ed_data$HRSWKGR, ed_data$JOBSATIS),1)
prop.table(table(ed_data$HRSWKGR, ed_data$JOBSATIS),2)
hour.prop <- prop.table(table(ed_data$JOBSATIS, ed_data$HRSWKGR))
mosaicplot(hour.prop, color=TRUE,main="Job satisfaction by hours per week worked")
```

Satisfaction verses Primary Work Activity
```{r}
ed_data <- ed_data %>% mutate(WAPRSM_non = ifelse(WAPRSM ==1, "Research",
                                    ifelse(WAPRSM == 2, "Teaching",
                                    ifelse(WAPRSM == 3,"Management", 
                                    ifelse(WAPRSM ==4, "Computers", "Other")))))


waprsm <- prop.table(table(ed_data$JOBSATIS, ed_data$WAPRSM_non),1)
waprsm
mosaicplot(waprsm, color=TRUE,main="Satisfaction vs Primary Work Activity")
```


Interaction of Available Benefits Compared to Insurnace
```{r}
library(gridExtra) 

ed_data <- ed_data %>% mutate(JOBPENS = ifelse(JOBPENS ==1, "Yes", "No"))
ed_data <- ed_data %>% mutate(JOBVAC = ifelse(JOBVAC ==1, "Yes", "No"))

pens.prop <- prop.table(table(ed_data$JOBINS, ed_data$JOBPENS))
vac.prop <- prop.table(table(ed_data$JOBINS, ed_data$JOBVAC))
proft.prop <- prop.table(table(ed_data$JOBINS, ed_data$JOBPROFT))
```



Interaction of Available Benefits Compared to Insurnace
```{r}
library(gridExtra) 

ed_data <- ed_data %>% mutate(JOBPENS = ifelse(JOBPENS ==1, "Yes", "No"))
ed_data <- ed_data %>% mutate(JOBVAC = ifelse(JOBVAC ==1, "Yes", "No"))

pens.prop <- prop.table(table(ed_data$JOBINS, ed_data$JOBPENS))
vac.prop <- prop.table(table(ed_data$JOBINS, ed_data$JOBVAC))
proft.prop <- prop.table(table(ed_data$JOBINS, ed_data$JOBPROFT))
```


Proportion of Insurance vs Pension Plan
```{r}
pens.prop
```

Proportion of Insurance vs Paid Vacation/Sick/Leave Days

```{r}
vac.prop
```

Proportion of Insurance vs Profit Plan

```{r}
proft.prop
```

Mosaic Plots
```{r}
m1<- mosaicplot(pens.prop, color=TRUE,main="Insurance vs Pension Plan")
m2 <- mosaicplot(vac.prop, color=TRUE,main="Insurance vs Paid Vacation/Sick/Leave")
m3 <- mosaicplot(proft.prop, color=TRUE,main="Insurance vs Profit Plan")
```




```{r, echo=FALSE}
pens1.prop <- prop.table(table(ed_data$JOBSATIS, ed_data$JOBPENS))
vac1.prop <- prop.table(table(ed_data$JOBSATIS, ed_data$JOBVAC))
proft1.prop <- prop.table(table(ed_data$JOBSATIS, ed_data$JOBPROFT))
ins1.prop <- prop.table(table(ed_data$JOBSATIS, ed_data$JOBINS))
```


Proportion of Job Satisfaction vs Pension Plan

```{r}
pens1.prop
```

Proportion of Job Satisfaction vs Vacation/Sick/Leave Days
```{r}
vac1.prop
```

Proportion of Job Satisfaction vs Profit Plan

```{r}
proft1.prop
```


Proportion of Job Satisfaction vs Insurance Plan

```{r}
ins1.prop
```

Satisfaction verses Available Benefits Mosaic Plots
```{r}
mosaicplot(pens.prop, color=TRUE,main="Satisfaction vs Pension Plan")
mosaicplot(vac.prop, color=TRUE,main="Satisfaction vs Paid Vacation/Sick/Leave")
mosaicplot(proft.prop, color=TRUE,main="Satisfaction vs Profit Sharing Plan")
mosaicplot(proft.prop, color=TRUE,main="Satisfaction vs Insurance Plan")
```




Proportion of Insurance vs Pension Plan
```{r}
pens.prop
```

Proportion of Insurance vs Paid Vacation/Sick/Leave Days
```{r}
vac.prop
```

Proportion of Insurance vs Profit Plan
```{r}
proft.prop
```


Salary
```{r}
library(plyr)
ed_data <- ed_data %>% mutate(JOBSATIS_cat = ifelse(JOBSATIS ==1, "Very Satisfied",
                                            ifelse(JOBSATIS ==2, "Somewhat Satisfied",
                                            ifelse(JOBSATIS ==3, "Somewhat Disatisfied","Very Disatisfied"))))


s_meds <- ddply(ed_data, .(JOBSATIS), summarise, med = median(SALARY))

ggplot(ed_data, aes(x=as.factor(JOBSATIS), y=SALARY)) + geom_boxplot() + xlab("Job Satisfaction") + ylab("Salary") +  geom_text(data = s_meds, aes(x = JOBSATIS, y = med, label = med), size = 3, vjust = -1.5) + ggtitle("Job Satisfaction vs Salary")

```







