---
title: "EDA"
output:
  word_document: default
  pdf_document: default
---

```{r, echo=FALSE, include = FALSE}
library(tidyverse)
library(mosaic)
library(readr)
library(ipumsr)
library(labelled)
library(dplyr)
library(ggplot2)

highered_00006 <- read_csv("/home/pelayo1/Stat_316_Project/highered_00006.csv")
View(highered_00006)
```

```{r, data cleaning}
usa_ddi <- read_ipums_ddi("/home/pelayo1/Stat_316_Project/highered_00006.xml")
ed_data <- read_ipums_micro(usa_ddi, verbose = FALSE)

ed_data <- ed_data %>% 
  filter(LFSTAT==1, JOBSATIS != 98) %>%
  mutate(GENDER = droplevels(as_factor(GENDER)), 
         RACETH= droplevels(as_factor(RACETH)), 
         DGRDG= droplevels(as_factor(DGRDG)), 
         LFSTAT= droplevels(as_factor(LFSTAT)), 
         HRSWKGR = droplevels(as_factor(HRSWKGR)), 
         JOBINS = droplevels(as_factor(JOBINS)),
         JOBPENS=droplevels.factor(JOBPENS),
         JOBPROFT=droplevels(as_factor(JOBPROFT),
        JOBVAC = droplevels(as_factor(JOBVAC)),
        OCEDRLP = droplevels(as_factor(OCEDRLP)), 
        NOCPRMG = droplevels(as_factor(NOCPRMG)),
        WAPRSM= droplevels(as_factor(WAPRSM)),
        SALARY = droplevels(as_factor(SALARY)), 
        JOBSATIS = droplevels(as_factor(JOBSATIS))))
ed_data <- ed_data %>% mutate(OCEDRLP = ifelse(OCEDRLP == 1,"Closely Related",
                                    ifelse(OCEDRLP == 2, "Somewhat Related",
                                    ifelse(OCEDRLP == 3, "Not Related",
                                    ifelse(OCEDRLP == 98, "Logical Skip", NA)))))
ed_data <- ed_data %>% mutate(NOCPRMG = ifelse(NOCPRMG == 1,"Computer and mathematical scientists ",
                                    ifelse(NOCPRMG == 2, "Biological, agricultural and other life scientists",
                                    ifelse(NOCPRMG == 3, "Physical and related scientists",
                                    ifelse(NOCPRMG == 4, "Social and related scientists",
                                    ifelse(NOCPRMG == 5, "Engineers",
                                    ifelse(NOCPRMG == 6, "Science and engineering related occupations ",
                                    ifelse(NOCPRMG == 7, "Non-science and engineering occupations ",
                                    ifelse(NOCPRMG == 98, "Logical Skip", NA)))))))))
head(ed_data)
```


```{r, EDA from Major Related}
#ggplot(ed_data, aes(GENDER, AGE)) + geom_boxplot() + ylab("Age") + xlab("Gender")

g <- ggplot(ed_data, aes(JOBSATIS, fill= OCEDRLP)) + geom_bar()
g
plyr::count(ed_data, "OCEDRLP")
table(ed_data$JOBSATIS, ed_data$OCEDRLP)

(30103+26764)/(30103+26764+3944+791) #Percent of people who are satisfied out of people who's jobs are closely related to their major

(384+5597)/(384+5597+1744+837) #Percent of people who are satisfied out of people who's jobs are not related to their major

(9170+12147)/(9170+12147+2490+590) #Percent of people who are satisfied out of people who's jobs are somewhat related to their major

table(ed_data$JOBSATIS, ed_data$OCEDRLP)

```

92% of people who have a job closely related to their major are very/somewhat satisfied. 

87.3% of people who have a job somewhat related to their major are very/somewhat satisfied. 

70% of people who have a job not related to their major are very/somewhat satisfied. 


```{r, Analysis NOCPRNG}
ed_data <- ed_data %>% mutate(NOCPRMG_non = ifelse(NOCPRMG =="Computer and mathematical scientists ", "MSCS",
                                    ifelse(NOCPRMG == "Biological, agricultural and other life scientists", "Bio",
                                    ifelse(NOCPRMG == "Physical and related scientists", "Physics",
                                    ifelse(NOCPRMG =="Social and related scientists", "Social",
                                    ifelse(NOCPRMG =="Engineers", "Engineers",
                                    ifelse(NOCPRMG =="Science and engineering related occupations ", "S-Engineer-related", "NS-Engineer-related")))))))


table(ed_data$NOCPRMG_non, ed_data$JOBSATIS)

prop.table(table(ed_data$NOCPRMG, ed_data$JOBSATIS),2) #Divides by Column
prop.table(table(ed_data$NOCPRMG, ed_data$JOBSATIS),1) #Divides by Row

ed_data %>% 
  count(NOCPRMG_non ,JOBSATIS) %>%  
  ggplot(mapping = aes(x = NOCPRMG, y = as.factor(JOBSATIS))) +
    geom_tile(mapping = aes(fill = n))

```

Out of all the people who are very satisfied, 30% of them are in Non-science and engineering occupations, and 21% are in Science and engineering related occupations. Out of all the people who are disatisfied, 49% of them are in non-science and engineering occupations. 

Out of the people in non-science and engineering occupations, 86% of them are very/somewhat satisfied. Out of the people in science and engineering related occupations, 91% of them are very/somewhat satisfied. 

```{r}
ed_data <- ed_data %>% mutate(OCEDRLP_non = ifelse(OCEDRLP =="Closely Related", "Close",
                                    ifelse(OCEDRLP == "Somewhat Related", "Somewhat", "Not")))

prop.table(table(ed_data$NOCPRMG_non, ed_data$OCEDRLP_non),2) #Divides by Column
prop.table(table(ed_data$NOCPRMG_non, ed_data$OCEDRLP_non),1)

mosaicplot(prop.table(table(ed_data$NOCPRMG_non, ed_data$OCEDRLP_non)), color=TRUE,main="Overall NOCPRMG vs OCEDRLP Proportions")

```

The field with the most people closely related to their major group is Science engineer related field with 23.2%, followed by Non-Science Engineer-Related people with 21.3%. The most people who are in a job of a not related field is in Non-Science engineering related jobs (66%). However, out of all the people who are in Non-Science and engineering related fields, 26% are not satisfied and 43% are satisfied. 


```{r}
library(gridExtra) 

ed_data <- ed_data %>% mutate(JOBPENS = ifelse(JOBPENS ==1, "Yes", "No"))
ed_data <- ed_data %>% mutate(JOBVAC = ifelse(JOBVAC ==1, "Yes", "No"))


pens.prop <- prop.table(table(ed_data$JOBINS, ed_data$JOBPENS))
vac.prop <- prop.table(table(ed_data$JOBINS, ed_data$JOBVAC))
proft.prop <- prop.table(table(ed_data$JOBINS, ed_data$JOBPROFT))

pens.prop
vac.prop
proft.prop

m1<- mosaicplot(pens.prop, color=TRUE,main="Insurance vs Pension Plan")
m2 <- mosaicplot(vac.prop, color=TRUE,main="Insurance vs Paid Vacation/Sick/Leave")
m3 <- mosaicplot(proft.prop, color=TRUE,main="Insurance vs Profit Plan")
```

72.2% of people who have an insurance plan also have a pension plan. 78.5% of people who have an insurance plan also have paid leave/sick/vacation days. However, only 23.3% of people with insurance plans also have a pension plan. Infact, it appears if someone does not have an insurance plan in their company, then 61% of them have a pension plan.

```{r}
ed_data <- ed_data %>% mutate(WAPRSM = ifelse(WAPRSM ==1, "Research and Development",
                                    ifelse(WAPRSM == 2, "Teaching",
                                    ifelse(WAPRSM == 3,"Management and Administration", 
                                    ifelse(NOCPRMG ==4, "Computer Applications", "Other")))))



```


```{r}
pens1.prop <- prop.table(table(ed_data$JOBSATIS, ed_data$JOBPENS))
vac1.prop <- prop.table(table(ed_data$JOBSATIS, ed_data$JOBVAC))
proft1.prop <- prop.table(table(ed_data$JOBSATIS, ed_data$JOBPROFT))
ins1.prop <- prop.table(table(ed_data$JOBSATIS, ed_data$JOBINS))

pens1.prop
vac1.prop
proft1.prop
ins1.prop


mosaicplot(pens.prop, color=TRUE,main="Satisfaction vs Pension Plan")
mosaicplot(vac.prop, color=TRUE,main="Satisfaction vs Paid Vacation/Sick/Leave")
mosaicplot(proft.prop, color=TRUE,main="Satisfaction vs Profit Sharing Plan")
mosaicplot(proft.prop, color=TRUE,main="Satisfaction vs Insurance Plan")


```

If the job has a pension plan 68% of people are either very satisfied or somewhat satisfied. 74% percent of people are either very satisfied or somewhat satisfied if the job has paid vacation/sick/leave days. It is interesting to see that 67% of people seem to still be somewhat satisfied or very satisfied even if the job does not have a profit sharing plan. 76% of people who have an insurance plan are either very satisfied or somewhat satisfied.


```{r}
library(plyr)
ed_data <- ed_data %>% mutate(JOBSATIS_cat = ifelse(JOBSATIS ==1, "Very Satisfied",
                                            ifelse(JOBSATIS ==2, "Somewhat Satisfied",
                                            ifelse(JOBSATIS ==3, "Somewhat Disatisfied","Very Disatisfied"))))


s_meds <- ddply(ed_data, .(JOBSATIS), summarise, med = median(SALARY))

ggplot(ed_data, aes(x=as.factor(JOBSATIS), y=SALARY)) + geom_boxplot() + xlab("Job Satisfaction") + ylab("Salary") +  geom_text(data = s_meds, aes(x = JOBSATIS, y = med, label = med), size = 3, vjust = -1.5)

```

As the median salary decreases, job satisfaction decreases as well. 

The majority of our sample are satisfied with their jobs. 
```{r, EDA from AGE}
ed_data %>%
  count(JOBSATIS)
ggplot(data = ed_data) + 
  geom_bar(mapping = aes(x = JOBSATIS))
```
There is a wide range of age group from mid-20s to late-70s. People who are very satisfied have slightly older age range. 
```{r}
#Age
favstats(~AGE, data=ed_data)
ggplot(data=ed_data) + 
  geom_histogram(mapping = aes(x=AGE))
favstats(AGE~JOBSATIS, data=ed_data)
ggplot(data = ed_data) + 
  geom_boxplot(mapping = aes(x = as.factor(JOBSATIS), y=AGE))
```
A higher proportion of male are very satisfied with their jobs compared to female; while a higher proportion of female are very dissatisfied with their jobs compared to male. 
```{r}
#Gender
prop.table(table(ed_data$GENDER))
prop.table(table(ed_data$GENDER, ed_data$JOBSATIS),1)
prop.table(table(ed_data$GENDER, ed_data$JOBSATIS),2)

gender.prop <- prop.table(table(ed_data$JOBSATIS, ed_data$GENDER))
mosaicplot(gender.prop, color=TRUE,main="Job satisfaction by gender")
```
The majority of our sample is White, and a higher proportion of White people are very satisfied with their jobs compared to Asian and under-represented minorities. a higher proportion of under-represented minorities are very dissatisfied with their jobs compared to White and Asian people. 
```{r}
#Race
prop.table(table(ed_data$RACETH))
prop.table(table(ed_data$RACETH, ed_data$JOBSATIS),1)
prop.table(table(ed_data$RACETH, ed_data$JOBSATIS),2)
race.prop <- prop.table(table(ed_data$JOBSATIS, ed_data$RACETH))
mosaicplot(race.prop, color=TRUE,main="Job satisfaction by race")
```
A high percentage of those who had a Professional degree (54.7%) were very satisfied with their job compared to people who had other degrees (40-47%). 
```{r}
#Highest degree
prop.table(table(ed_data$DGRDG))
prop.table(table(ed_data$DGRDG, ed_data$JOBSATIS),1)
prop.table(table(ed_data$DGRDG, ed_data$JOBSATIS),2)
degree.prop <- prop.table(table(ed_data$JOBSATIS, ed_data$DGRDG))
mosaicplot(degree.prop, color=TRUE,main="Job satisfaction by highest degree")
```
The majority works more than 36 hours per week, and the proportion of those who work more than 40 hours/week that are very dissatisfied with their jobs is smaller compared to those who work less than 40 hours/week. 
```{r}
#Hours/week worked
prop.table(table(ed_data$HRSWKGR))
prop.table(table(ed_data$HRSWKGR, ed_data$JOBSATIS),1)
prop.table(table(ed_data$HRSWKGR, ed_data$JOBSATIS),2)
hour.prop <- prop.table(table(ed_data$JOBSATIS, ed_data$HRSWKGR))
mosaicplot(hour.prop, color=TRUE,main="Job satisfaction by hours per week worked")
```

